Follow-up Samples

### 2017-02-07 submission/manuscript_outline_20161024.Rmd

**Background:** Colorectal cancer (CRC) continues to be a worldwide health problem with early detection being used as a key component in mitigating deaths due to the disease. Previous research suggests a link between fecal bacterial microbiome and CRC.   The overall objective of our study was to investigate the changes in the bacterial microbiome after surgery in patients with lesion (i.e. adenoma or carcinoma).  Specifically, we wanted to identify what within the community was different within those undergoing surgical removal of lesion.  We also wanted to investigate the use of the bacterial microbiome and Fecal Immunoglobulin Test (FIT) to build models which could classify individuals as having a lesion or as before surgery.     
**Results:** Adenoma individual's bacterial microbiome were more similar to their pre-surgery sample then those with carcinoma (P-value = `r format(change_summary['thetayc', 'Pvalue'], digits = 3)`).  Their change in FIT was also significantly different then those with carcinoma (P-value = `r format(change_summary['fit', 'Pvalue'], scientific = TRUE, digits = 3)`).  There was no significant difference in any individual OTU between samples before and after surgery (P-value > `r format(min(filter(all_OTU_pair_wilcox, analysis == "ALL")[, 'BH_corrected']), digits = 3)`).  A model with a total of `r length(lesion_top10_varsMDA$variable)` variables was able to classify lesion (AUC = `r format(max(reduced_lesion_AUC_run_summary$ROC), digits = 3)` - `r format(min(reduced_lesion_AUC_run_summary$ROC), digits = 3)`) while the model to classify samples as before surgery had `r length(IF_top10_varsMDA$variable)` variables (AUC = `r format(max(reduced_IF_AUC_run_summary$ROC), digits = 3)` - `r format(min(reduced_IF_AUC_run_summary$ROC), digits = 3)`).  There was a significant decrease in the model positive probability for the after surgery sample being either a lesion or before surgery sample (P-value = `r format(filter(all_probs_paired_wilcox, model_type == "red_lesion" & comparison =="lesion")[, "BH_correction"], scientific = TRUE, digits = 3)` and `r format(filter(all_probs_paired_wilcox, model_type == "red_IF" & comparison =="lesion")[, "BH_correction"], scientific = TRUE, digits = 3)`).  In total there were `r length(common_vars_summary$otu)` OTUs that were common to both models and the majority of these classified to commensal bacteria (`r common_vars_summary["Otu000005", "lowest_ID"]`, `r common_vars_summary["Otu000012", "lowest_ID"]`, `r common_vars_summary["Otu000082", "lowest_ID"]`, and `r paste(gsub("_unclassified", "", common_vars_summary["Otu000006", "lowest_ID"]))`).   
**Conclusions:** Our data suggests that treatment not only significantly reduces the probability of having a colonic lesion within our models but also causes detectable changes in the bacterial microbiome.  Further surveillance of these individuals will enable us to determine whether models such as the ones we present could be used to predict recurrence of colorectal cancer.

@@

bacterial microbiome; colorectal cancer; polyps; FIT; detection; risk factors

>>>>

**Background:** Colorectal cancer (CRC) continues to be a worldwide health problem with previous research suggesting that a link may exist between the fecal bacterial microbiome and CRC.  The overall objective of our study was to test the hypothesis that changes in the bacterial microbiome occur after surgery in patients with lesions (i.e. adenoma or carcinoma).  Specifically, we wanted to identify what within the community was different before and after surgical removal of said lesion.

**Results:** The bacterial microbiome in the pre and post surgery samples of the individuals with adenoma were similar while those with carcinoma were not (P-value = `r format(change_summary['thetayc', 'Pvalue'], digits = 3)`).  There was no difference in the relative abundance of any OTU between the pre and post surgery samples (P-value > `r format(min(filter(all_OTU_pair_wilcox, analysis == "ALL")[, 'BH_corrected']), digits = 3)`).  A model with a total of `r length(lesion_top10_varsMDA$variable)` variables was able to classify lesion (AUC = `r format(max(reduced_lesion_AUC_run_summary$ROC), digits = 3)` - `r format(min(reduced_lesion_AUC_run_summary$ROC), digits = 3)`) while a model built to classify samples as before surgery had `r length(IF_top10_varsMDA$variable)` variables (AUC = `r format(max(reduced_IF_AUC_run_summary$ROC), digits = 3)` - `r format(min(reduced_IF_AUC_run_summary$ROC), digits = 3)`).  The post surgery sample for both models had a decrease in the positive probability for either lesion or before surgery sample (P-value = `r format(filter(all_probs_paired_wilcox, model_type == "red_lesion" & comparison =="lesion")[, "BH_correction"], scientific = TRUE, digits = 3)` and `r format(filter(all_probs_paired_wilcox, model_type == "red_IF" & comparison =="lesion")[, "BH_correction"], scientific = TRUE, digits = 3)`).  In total there were `r length(common_vars_summary$otu)` OTUs that were common to both models and the majority of these classified to commensal bacteria (`r common_vars_summary["Otu000005", "lowest_ID"]`, `r common_vars_summary["Otu000012", "lowest_ID"]`, `r common_vars_summary["Otu000082", "lowest_ID"]`, and `r paste(gsub("_unclassified", "", common_vars_summary["Otu000006", "lowest_ID"]))`).   
**Conclusions:** Our data supports the hypothesis that there are differences in the bacterial microbiome between pre and post surgical samples.  With individuals with carcinoma having more drastic changes to the overall community then those with adenoma.  However, changes to commensal bacteria were common to individuals with either adenoma or carcinoma suggesting that these bacteria may be central to initial polyp formation and transition to carcinoma.

@@

bacterial microbiome; colorectal cancer; polyps; FIT; post surgery; risk factors
